---
title: "Main Functions"
format: html
---

## load_data()

```{r}
plastics <- load_data()
```

This loads the full merged plastic waste dataset from the package, combining
the TidyTuesday BFFP audit data with country-level ecology and population data.

## filter_country()

```{r}
filter_country(plastics, "India", positive_only = TRUE)
```

This filters to return a data frame with just the Grand Total column for a 
specific country. There is an option to return values where Grand Total is all
positive values.

## plastics_long()

```{r}
plastics_long(plastics)
```

This transposes a wide format data frame into long format, while also adding a
user-friendly column showing plastic type names.

## clean_plastics()

```{r}
plastics <- clean_plastics(plastics)
```

This standardizes inconsistent country names (e.g. recoding `"Korea"` to
`"South Korea"`, removing junk rows), and adds plastic proportion columns
for each plastic type as a share of grand total.

## summarize_country()

```{r}
summarize_country(plastics, "India")
```

This returns a one-row summary of a country's average plastic type
proportions across all years, along with its dominant plastic type.

## summary_category_metrics()

```{r}
summary_category_metrics(plastics, env_var = "threatened_species")
```

This bins countries into Low, Medium, and High tiers by an environmental
indicator and summarizes average plastic composition within each tier.

## plot_plastics()

```{r}
plot_plastics(plastics, country = "India", n = 3)
```

This produces a horizontal bar chart of a country's top plastic types,
labelled by their share of total plastic collected.

## plot_plastic_ridges()

```{r}
plot_plastic_ridges(plastics, plastics = c("pet", "hdpe", "pp"),
                   env_var = "threatened_species")
```

This draws weighted density ridges of an environmental indicator across
selected plastic types, with each country weighted by quantity collected.

## model_pet_dominant()

```{r}
model_pet_dominant(plastics)
```

This fits a logistic regression predicting whether PET is a country's
dominant plastic type from forested area and threatened species count,
returning a tidy table of coefficients and p-values.
